\chapter{OntoBG: An ontology of boardgames}

\todo[inline]{Esse texto está precisando de uma boa revisão, eu achei muito difícil de manter a leitura. Tá no nível C, sendo A inglês nativo, B inglês esperado por um não nativo e C tentando atingir o necessário, eu vou sair mudando o que já conseguir resolver. Já reformulei o próximo parágrafo}

This chapter covers the OntoBG ontology in detail: its specification, diagrams and restrictions, and also reviewing some important conceptual points of the model. It also provides insights on how to use it for different modeling purposes.

% se repete
%OntoBG, by what this works propose, is defined and explained here. 
Following the scope delimitation of this work it does not cover the whole possibilities of board games. It aims to provide new insight on board game modeling and study, not thoroughly explain the domain.

The ontology is represented as four diagrams. One that provides an overview of the MDA structure throughout the ontology, and three others that convey the understanding of each part of the MDA, which are: OntoBG-M, OntoBG-D and OntoBG-A. The following sections, present and explain each of them. At the end of the chapter, an overall perspectivee of the ontology is given. 
%The last three diagrams are too big to fit inside the text and are present in [APENDIX X-> X+2].
\todo{Colocar diagramas dos 3 modelos.}


Before understanding the model, there is one more concept to be created: the relationship of cause and consequence. Its need is fully explained in the next section.

\input{Contents/cap5.ontology/causalrelation.tex}
\input{Contents/cap5.ontology/mda.tex}
\input{Contents/cap5.ontology/mechanics.tex}
\input{Contents/cap5.ontology/dynamics.tex}
\input{Contents/cap5.ontology/aesthetics.tex}

