\chapter{Crafting an Ontology}

This chapter covers how a domain ontology about boardgames was built. The ontology, henceforth called OntoBG, is composed of 3 blocks and constructed upon the theoretical framework of the Universal Foundational Ontology or UFO \citep{guizzardi_ontological_2005}, using the OntoUML modeling language as described in \cite{guizzardi_ontological_2005,guizzardi_ontoUML_2004,guizzardi2015towards}. 

The purpose of separating the ontology into three sub-ontologies is to put together the three ontologies of mechanics, dynamics and aesthetics in a coherent and understandable single ontology, this will expose the structure of the MDA framework used as theoretical principle of this thesis. Each of those sub-ontologies is assembled using a specific methodology. The particulars of their creation will be further explained in the following sections.

For clarity purposes, these three sub-ontologies of the ontology are named OntoBG-M, OntoBG-D and OntoBG-A, respectively for the mechanics, dynamics and aesthetics.

Before introducing the specifics for each sub-ontology of the OntoBG it is important to specify how it can be used to model this particular domain. Due to the broad philosophical foundation of UFO, this subject will be split in two sections. The first addressing this philosophical theory and the general concepts in this model, afterwards it is addressed how it will be applied in the domain of boardgames to give fruition to OntoBG.


\input{Contents/cap3.methodology/ThinkingUFO.tex}
\input{Contents/cap3.methodology/BoardgameModel.tex}

\input{Contents/cap3.methodology/Mechanics.tex}
\input{Contents/cap3.methodology/Dynamics.tex}
\input{Contents/cap3.methodology/Aesthetics.tex}
\input{Contents/cap3.methodology/Usability.tex}
\input{Contents/cap3.methodology/Scope.tex}


