\section{Concepts}
In this section some important concepts used in this work are exposed and explained in accord to their meaning to this dissertation. To be clear about what they mean whenever they are used. As it is not in the scope of this work there will be no rigid definition of the word Game. When looking into the definitions given by the many authors mentioned in section 1.1 it is possible to understand what is a game to this work. 

\textbf{Boardgame} is used with the idea of a non-digital game. It does not need to have a board, it can be a game made of only cards or even a mimic game with no object, only the rules. The scope of this work does not include digital games. We use the word boardgame, instead of game, throughout this dissertation to emphasize this.

The terms, \textbf{mechanics}, \textbf{dynamics} and \textbf{aesthetics} are used with accord to their definition in the MDA framework, which is explained in the next chapter. 

\textbf{Magic Circle} is a notion first introduced in \cite{huizinga2014homo}. Although \citeauthor{huizinga2014homo} did not named it so, many scholars that further developed the concept name it as Magic Circle. \cite{salen2004rules,bateman_implicit_2015} are only a few examples. This concept develops on the idea that there is a certain boundary in reality in which a play of a game takes place and this boundary prohibits some aspects of reality to affect the game space, in other words it protects the game from the real world. The Magic Circle became very important to the studies of games and whenever mentioned throughout this work the Magic Circle significance is in accordance to the understanding given by \cite{salen2004rules}.

\textbf{World} is addressed in the modeling concept of the real world. A given world is a possible reality of facts and beings that are part of the modeled domain. In other words a world is a possible existence. An example, say that a model of families exist. One world is that Carlos, Mandy and Bob form a family while Eduardo, Sandra and Charlie form another. A second possible world is that Carlos, Eduardo and Bob from the first family while the others form the second one. 

