\section{Why games?}

Nowadays we live surrounded by games. There are games in smart-phones that are played while we come and go to our everyday tasks, games in our houses, in computers and even where there seems to be no games, there are people talking about games.

Financially, games are a big industry. The revenue of many video game companies are in the score of billions of dollars. \citet{babb2013comparing} evaluates diverse companies and games, giving a good illustration of this big industry. Only the US video game market is worth over 20 billion dollars. A single game, World of Warcraft, had almost 1 billion dollars revenue by itself. All of this before 2011, when today this market has only increased. This shows how big games are in the modern world, which enforces why we should study them deeply.

Games are also becoming present on education. Game-based learning became a very popular approach to teaching in many levels of education. In \citet{hainey2016systematic} the sheer amount of practical experiences found is enough to show that games are a part of education in the twentieth century.

One can guess that, if games are so entwined in our daily lives, there will be a great understanding of them, of what games are, how games are made, what means to play a game. Surprisingly that is not true. Studies in games have a great amount of disagreement. Many aspects of games are met with different opinions among scholars. To begin with, there is no commonly accepted definition of the term game. 

Many researchers addressed the problem of finding a definition \citep{jarvinen2009games,salen2004rules,crawford1984art,schell2014art,juul2010game} but they could not agree on a single definition. Each of them pursued very different paths in their search for understanding, thus each defined games in their own way. Although there is some common ground between some of those definitions. \cite{salen2004rules} systematically compared most definitions from previous works and concluded there was no true agreement between them. \cite{wittgenstein_philosophical_2009} went further to state that there is no possible definition, ``game is a concept with blurred edges''. Such is the hardship of finding a definition for games.

Other aspects of games share this multitude of opinions. Game design or how to make games is a subject that presents a variety of ideas and models for its understanding. Although there are more common parts and shared ideas through many researches this subject is far from a global complete knowledge. There are many other questions pertaining games that remain without a completely accepted answer. Such as `why do people play games?', `how do people react to games?', `how do people behave in games?'.

In light of such confusion and dissent. The purpose of this work is to extend current understanding of what games are made of, of which parts compose this complex and intriguing whole. Hopefully, this will bring us closer to completely understanding games. At least it should provide a tool for researchers and designers that wish to adventure in the games domain to start to understand what it is. Therefore, this is an attempt to create a boardgame ontology using the ontological principles of UFO \citep{guizzardi_ontological_2005}. It will provide the foundation for a human oriented ontology. This ontology is structured upon the MDA framework \citep{Hunicke2004} as it is simple and widely accepted. 

This dissertations focus its efforts in boardgames, this is due the author's experience, but also because the boardgame industry is increasing rapidly over the past decade and becoming a very expressive part of the daily lives in many countries. In a market report, done by Grand View Research \footnote{\url{https://www.grandviewresearch.com/industry-analysis/playing-cards-board-games-market} Accessed in 12/03/2020}, the boardgame market size was almost 12 billion dollars, with an expected increase of 8,79\% over the next six years. It still does not compete with video games, but this growth implies that boardgames are becoming more and more important over the years and must be addressed accordingly.